Sulfur Dioxide on Water's Surface

### 2011-04-05 Conclusions.tex
Draft 2

The surface behavior of the \suldiox~molecules was also determined. Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry, so the spectra of the surface \suldiox~vibrational modes are still unreported. However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. This was previously reported by Dang.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.

>>>>

The surface behavior of the \suldiox~molecules was also determined. Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry, so the spectra of the surface \suldiox~vibrational modes are still unreported. However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. This was previously reported by Baer et al.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.

### 2011-04-05 DensityResults.tex
Draft 2

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation. However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

>>>>

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation. However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Baer et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

### 2011-04-05 Introduction.tex
Draft 2

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas. A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations. That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.

>>>>

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation. Atmospheric dust particles and gases have also been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas. A computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations. That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.

### 2011-04-05 Methods.tex
Draft 2

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Dang, et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}

>>>>

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}
